Human Factors Influencing Visual Statistical Inference

### 2013-06-21 sociology_paper.tex
cleaned up paragraph on demographics

Effect of graduate degree is estimated as 0.182 which is highly significant. For an 18-25 year old female in the United States, the expected difference in the probability of correct response, if she has a graduate degree as compared to a high school degree, is, for a lineup with random effect of zero:
$$ \frac {\exp(-0.683+0.182)}{1+\exp(-0.683+0.182)}-  \frac {\exp(-0.683)}{1+\exp(-0.683)} = 0.377- 0.336 =0.041.$$ 
For a lineup with a predicted random effect of one standard deviation, i.e. we move away one standard deviation of lineup variation by adding 2.293 to each of the exponent terms in the equation, this difference reduces to 0.024. For two standard deviations it further reduces to 0.003 which is practically very small. Figure~\ref{fig:practical_impact_graduate} illustrates this example showing the proportions of correct responses due to graduate degree as compared to high school degree. The highest practical difference in proportions of correct response is estimated as 0.045 which is very negligible. The difference further diminishes if we move one or two standard deviation of lineup variability. A similar calculation shows the same negligible impact of age level 36-40 (0.0533 at most)  and country (0.0424 at most) on the probability of correct response. Thus even though some of the factor levels are statistically significant, their practical impact is negligible.

>>>>

While some of the demographic factors are strongly statistically significant, the main source of variation in proportion correct is the lineup difficulty. For example, let's examine the effect of graduate degree. To see just how large the effect is, we examine the change in proportion correct for a (hypothetical) 18-25 year old female in the United States, with a graduate degree as compared to a high school degree, for an average difficulty lineup (random effect = 0). Plugging in the relevant quantities to the fitted model gives a difference equal to:
$$ \frac {\exp(-0.683+0.182)}{1+\exp(-0.683+0.182)}-  \frac {\exp(-0.683)}{1+\exp(-0.683)} = 0.377- 0.336 =0.041.$$ 
The person with a high school education on average picks the data plot in 33.5\% of lineups having average difficulty, as compare to 37.6\% if they have a graduate degree. This difference is reduced to 2\% for a lineup with one standard deviation order of magnitude difference in difficulty. For two standard deviations it further reduces to 0.3\%. Thus although there is statstically significant difference in proportion correct for some demographic factors, these are not practically significant differences.  Figure~\ref{fig:practical_impact_graduate} illustrates this example showing fitted models for a US 18-25 female with either a high school education or a graduate degree.  Similar calculations show the same negligible impact of age level 36-40 (0.0533 at most)  and country (0.0424 at most) on the probability of correct response. Thus even though some of the demographic factors are statistically significant, practically, demographics do not substantially influence the results.
